Enforced Invariant Confluence

### 2018-02-23 doc/paper_redux/iconfluence.tex
Finsihed new iconfluence definitions.

\begin{claim}\clmlabel{irec_implies_reachable}
  $\Irec{e} \implies \reachable{e}$
\end{claim}
\begin{elidableproof}
  We perform a structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case $s_0$.}
      Axiomatically, $\reachable{s_0}$.

@@

    \item
      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \implies I{e_1 \join e_2}$
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  First, we show that (1) implies (2). Axiomatically, $\reachable{s_0}$, so by
  (1), $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary expressions such that
  $\Irec{e_1}$ and $\Irec{e_2}$. By \clmref{irec_implies_reachable},
  $\reachable{e_1}$ and $\reachable{e_2}$. Thus, $\reachable{e_1 \join e_2}$,
  so by (1), $I(e_1 \join e_2)$.

@@

The notion of operation-based objects is also taken
from~\cite{shapiro2011conflict}. Note that we do not have a join function like
we did with state-based objects.

@@

The definition of invariants remains unchanged.

\begin{example}
  $(\nats, 0)$ is a distributed object. Let $f_k(x) = x + k$. $t(x) = f_x$ is a
  transaction. $I(x) = x \geq 0$ is an invariant.
\end{example}

@@

\subsubsection{Process-Based}
As with the state-based approach, the process-based model is most similar to
the process model described by Shapiro et al.\ in~\cite{shapiro2011conflict}.
We are given a set $\seq{p}{1}{n}$ of $n$ processors each of which begins with
state $s_0$. Repeatedly, a processor executes a transaction $t$ on its current
state $s$ and then transitions from state $s$ to state $t(s)(s)$. It also
broadcasts $t(s)$ to all other processes exactly once. When a processor with
state $s'$ receives $t(s)$, it transitions from state $s'$ to state $t(s)(s')$.

Consider any two processors $p_i$ and $p_j$ with states $s_i^0, \ldots, s_i^n$
and $s_j^0, \ldots, s_j^m$ such that every state $s_i^0, \ldots, s_i^n, s_j^0,
\ldots, s_j^m$ satisfies $I$. $T$ is invariant-confluent with respect to $I$,
abbreviated \defword{\Iconfluent{}}, if for all $t \in T$, $I(t(s_i^n)(s_i^n))$
implies $I(t(s_i^n)(s_j^m))$.

\subsubsection{Graph-Based}
In the graph-based model, we are given a directed acyclic graph in which each
vertex is associated with a state $s$ and each edge is labelled with a shadow
operation $t(s)$ where $s$ is associated with some vertex in the graph.
%
Consider two invariant-satisfying states $s$ and $s'$ with invariant satisfying
ancestors. $T$ is \defword{\Iconfluent{}} if for all $t$, $I(t(s)(s))$ implies
$I(t(s)(s'))$.

\subsubsection{Expression-Based}
The expression-based approach deals with expressions built from $s_0$ and
transactions in $T$. That is, expressions $e$ are built from the grammar
\[
  e ::= s_0 \mid t(e_1)(e_2)
\]
where $t$ corresponds to a transaction in $T$. We can evaluate an expression,
denoted $eval(e)$, in the obvious way:
\begin{mathpar}
  eval(s_0) \defeq s_0

@@

@@

\begin{definition}
  $T$ is \defword{\Iconfluent}, if for all expressions $e_1$ and $e_2$,
  $\Irec{e_1}$, $\Irec{e_2}$, and $I(t(e_1)(e_1))$ implies $I(t(e_1)(e_2))$.
\end{definition}

\todo{Define ``reachable'' states and describe that \Iconfluence{} is
equivalent to saying that reachable states are a subset of invariant satisfying
states.}
\todo{Maybe define Iconfluence to say that the initial state satisfies the
invariant?}

>>>>

\begin{claim}\clmlabel{state_irec_implies_reachable}
  $\Irec{e} \implies \reachable{e}$
\end{claim}
\begin{elidableproof}
  We perform structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case $s_0$.}
      Axiomatically, $\reachable{s_0}$.

@@

    \item
      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \implies I{e_1 \join e_2}$
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  First, we show that (1) implies (2). Axiomatically, $\reachable{s_0}$, so by
  (1), $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary expressions such that
  $\Irec{e_1}$ and $\Irec{e_2}$. By \clmref{state_irec_implies_reachable},
  $\reachable{e_1}$ and $\reachable{e_2}$. Thus, $\reachable{e_1 \join e_2}$,
  so by (1), $I(e_1 \join e_2)$.

@@

Like state-based objects, the notion of operation-based objects is taken
from~\cite{shapiro2011conflict}. Note that we do not have a join function like
we did with state-based objects.

@@

The definition of an invariant is the same in the state-based and
operation-based model.

\begin{example}
  $(\nats, 0)$ is a distributed object. $t(x) = y \mapsto x + y$ is a
  transaction that given a state $x$ returns a function $y \mapsto x + y$ that
  adds $x$ to its argument. $\setst{x \geq 0}{x \in \nats}$ is an invariant.
\end{example}

@@

\subsubsection{Process-Based}
As with the state-based approach, the process-based model is most similar to
the process model described by Shapiro et al.\ in~\cite{shapiro2011conflict}.
We are given a set $\seq{p}{1}{n}$ of $n$ processors each of which begins with
state $s_0$. A processor $p_i$ can do one of two things.

First, $p_i$ can execute a transaction $t$ on its current state $s$ and then
transitions from state $s$ to state $t(s)(s)$, given that $I(t(s)(s))$. If
$\lnot I(t(s)(s))$, then $p_i$ will not perform $t$. If $p_i$ does execute $t$,
then it broadcasts $t(s)$ to all other processors exactly once.

Second, $p_i$ can receive a broadcasted shadow transaction $t(s_j)$ from some
other processor $p_j$. When $p_i$ receives $t(s_j)$, it transitions from its
state $s_i$ to state $t(s_j)(s_i)$. When $p_i$ receives a shadow transaction,
it must execute it, even if $\lnot I(t(s_j)(s_i))$.

$T$ is \Iconfluent{} if every reachable state (including $s_0$) satisfies the
invariant.

\subsubsection{Graph-Based}
In the graph-based model, we are given a directed acyclic graph in which each
vertex is a state $s$ and each edge is labelled with a shadow operation $t(s)$
where $s$ is some other vertex in the graph. $T$ is \Iconfluent{} if every
vertex in the graph satisfies $I$.

\subsubsection{Expression-Based}
Given an operation-based object $(S, s_0)$ and a set of transactions $T$, we
consider expressions built from the following grammar:
\[
  e ::= s_0 \mid t(e_1)(e_2)
\]
where $t$ corresponds to a transaction in $T$. We can evaluate an expression,
denoted $eval(e)$, in the obvious way:
\begin{mathpar}
  eval(s_0) \defeq s_0

@@

\begin{definition}
  As with state-based expressions, we formalize which operation-based
  expressions are reachable.

    \inferrule{
      \reachable{e_1} \\
      \reachable{e_2} \\
      I(t(e_1)(e_1))
    }{\reachable{t(e_1)(e_2)}}
  \end{mathpar}
\end{definition}

\begin{definition}
  $T$ is \Iconfluent{} if $\setst{e}{\reachable{e}} \subseteq I$.
\end{definition}

As with state-based objects, we have an equivalent definition of \Iconfluence{}
that deals with the closure of recursively invariant satisfying states.

@@

\begin{claim}\clmlabel{op_irec_implies_reachable}
  $\Irec{e} \implies \reachable{e}$
\end{claim}
\begin{elidableproof}
  We perform a structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case $s_0$.}
      Axiomatically, $\reachable{s_0}$.

    \item \textbf{Case $t(e_1)(e_2)$.}
      $\Irec{t(e_1)(e_2)}$, so $\Irec{e_1}$, $\Irec{e_2}$, and $t(e_1)(e_1)$.
      By the inductive hypothesis, $\reachable{e_1}$ and $\reachable{e_2}$.
      Thus, $\reachable{t(e_1)(e_2)}$.
  \end{itemize}
\end{elidableproof}

\begin{claim}
  Given an operation-based object $(S, s_0)$, a set of transactions $T$, and an
  invariant $I$, the following two are equivalent:
  \begin{enumerate}[\quad(1)]
    \item
      $\setst{e}{\reachable{e}} \subseteq I$

    \item
      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \land I(t(e_1)(e_1)) \implies
      I(t(e_1)(e_2))$.
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  First, we show that (1) implies (2). $\reachable{s_0}$, so by (1), $I(s_0)$.
  Let $e_1$ and $e_2$ be arbitrary expressions such that $\Irec{e_1}$,
  $\Irec{e_2}$, and $I(t(e_1)(e_1))$. By \clmref{op_irec_implies_reachable},
  $\reachable{e_1}$ and $\reachable{e_2}$, so $\reachable{t(e_1)(e_2)}$. By
  (1), $I(t(e_1)(e_2))$.

  Next, we show that (2) implies (1). We show by structural induction that
  $\reachable{e} \implies \Irec{e}$. From this, (1) is immediate.
  \begin{itemize}
    \item \textbf{Case $s_0$.}
      By (2), $I(s_0)$, so $\Irec{s_0}$.

    \item \textbf{Case $t(e_1)(e_2)$.}
      $\reachable{t(e_1)(e_2)}$, so $\reachable{e_1}$, $\reachable{e_2}$, and
      $I(t(e_1)(e_1))$. By the inductive hypothesis, $\Irec{e_1}$ and
      $\Irec{e_2}$. By (2), $I(t(e_1)(e_2))$, so $\Irec{t(e_1)(e_2)}$.
  \end{itemize}
\end{elidableproof}
